---
title: "Assignment 1"
date: "2022-11-13"
---
Our dataset is a list of startups getting investments from a variety of domestic and international investors. Data is gathered from KPMG and 212’s [Turkish Startup Investments Review 2021 report](https://assets.kpmg/content/dam/kpmg/tr/pdf/2022/03/turkish-startup-investments-review-2021.pdf).

[Startup Deals 2021 Data](https://github.com/berkorbay/datasets/raw/master/startup_deals/startup_deals_2021.xlsx)


## Key Takeaways

- Top 5 sectors that are invested most in terms of total value are **Ecommerce enabler, Delivery and Logistics, Gaming, SaaS and Fintech** respectively.

- Most of the investment deals are sealed during period from **June 2021 to August 2021**

- Regarding the distribution of investments made to the target companies, first 4 startups, namely **Trendyol, Getir, Hepsiburada and Dream Games**,  attracted a total of 3.4 billion USD approximately and accounted for **89.2% of all investments made in 2021**.

- The share of **local investors**, with a total investment value of **273 million USD**, is limited and majority of the investments are made by **foreign investors** with a total value of **3.5 billion USD**. 

- Most of the investors by far the most are from **Turkey** in terms of *number of deals*, where as **United States, UAE and Qatar** are leading countries of investors in terms of *value* invested in startups.

## Data Wrangling and Exploratory Data Analysis

Download necessary libraries

```{r,output=FALSE, warning=FALSE}

library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(scales)
library(knitr)
library(viridis)
library(maps)
```

### Data Preparation

Necessary data manipulation operations are made to prepare raw data for further analysis.

```{r data_prep, output = FALSE, warning=FALSE}

#read data
startup <- read_excel("startup_deals_2021.xlsx")

# Convert spaces to underscore in column names
colnames(startup) <- startup %>% 
  colnames() %>% 
  gsub(" ", "_", .)

# Convert data type of numeric columns from string to numeric
startup$`Deal_Value_(USD)` <- as.numeric(startup$`Deal_Value_(USD)`)

# Convert percentages to numeric format
startup$`Stake_(%)` <- startup$`Stake_(%)` %>% 
  gsub("\\..*", "", .) %>% 
  gsub("%", "", .) %>% 
  as.numeric()

# Convert string date to date format
startup$Announcement_Date <- startup$Announcement_Date %>% 
  paste("01", ., sep = "") %>% 
  parse_date_time(., "dmy")

# New columns for months and years
startup$year <- year(startup$Announcement_Date)
startup$month <- month(startup$Announcement_Date)
str(startup)

```

### Determining Top 5 Sectors in terms of Total Value of Investments

Table below shows the top 5 sectors of startups that are invested most in total value in 2021. 

```{r, warning=FALSE}

top_5_sector_table <- startup %>% 
  group_by(Sector) %>% 
  summarize(total_value = sum(`Deal_Value_(USD)`, na.rm = TRUE)) %>% 
  arrange(desc(total_value)) %>% 
  top_n(n = 5)

kable(top_5_sector_table)

```

### Visualising Monthly Sectoral Investments to Startups in 2021

Sectors are grouped as top 5 sectors and the remaining sectors are labeled as **"Others"**. As can be seen from the bar plot of monthly distribution of value of total investments, highest value of deals are sealed and announced during summer in 2021.

```{r, warning=FALSE}

# Obtaining top 5 sectors found in previous code as a vector.
top_5_sector <- top_5_sector_table %>% 
  select(Sector) %>% 
  pull()

# Create a new variable that classify sectors other than top_5 as others
startup <- startup %>% 
  mutate(group_sector = case_when(Sector %in% top_5_sector ~ Sector,!(Sector %in% top_5_sector) ~ "Other"))

# Plotting monthly investment values of top_5_sectors and others
startup %>% 
  group_by(Announcement_Date, group_sector) %>% 
  summarise(value = sum(`Deal_Value_(USD)`, na.rm = TRUE)) %>% 
  ggplot(aes(x = Announcement_Date, y = value, fill = group_sector)) +
  geom_col() +
  scale_y_continuous(labels = comma_format(big.mark = ".",
                                           decimal.mark = ",")) +
  labs(x = "Deal Announcement Date", y = "Total Deal Value (USD)") + 
  scale_fill_viridis(discrete=TRUE) +
  theme_bw()
```

### Distribution of Investments Across Target Companies.

The deals accomplished in 2021 is concentrated on a few startups. 4 highest valued deals sum up to almost 90% of total value of all deals in 2021.

```{r, warning=FALSE}
# Obtain percentages of Target_companies in terms of value invested in 2021
# Classify companies as first 4 companies and others

startup_percent <- startup %>% 
  group_by(Target_Company) %>% 
  summarize(Total_Deal_Value = sum(`Deal_Value_(USD)`, na.rm = TRUE)) %>% 
  mutate(percent = Total_Deal_Value / sum(Total_Deal_Value) * 100) %>% 
  arrange(desc(percent)) %>% 
  mutate(group_company = case_when(Target_Company %in% Target_Company[1:4] ~ Target_Company, !(Target_Company %in% Target_Company[1:4])~"others"))

startup_percent[1:4,] %>% 
  summarize(Total_Value = sum(Total_Deal_Value), percentage = sum(percent))

# Draw a column chart according to group framework given above

startup_percent_group <- startup_percent %>% 
  group_by(group_company) %>% 
  summarize(group_percent = sum(percent, na.rm = TRUE)) %>% 
  arrange(desc(group_percent)) %>% 
  mutate(cum_percentage = cumsum(group_percent)) 

kable(startup_percent_group)

startup_percent_group %>% 
  arrange(group_percent) %>% 
  mutate(group_company = factor(group_company, group_company)) %>% 
  ggplot(aes(x = group_company, y = group_percent)) +
  geom_col(fill = "#69b3a2") + 
  labs(x = "Target Company", y = "Percentage (%)", title = "Distribution of Startup Investments among Target Companies in 2021") +
  theme_bw()
  
```

### Categorizing Investors as Locals and Foreigners

Equal partnership in investments is assumed if number of investors is higher than one in a particular deal. 

```{r, warning=FALSE}
# Generate Unique id for each transaction and calculate number of investors as a new column "count"
id_count <- startup %>% 
  mutate(id = 1:dim(startup)[1]) %>% 
  separate_rows(`Investor's_Origin`, sep = ", ") %>% 
  group_by(id) %>% 
  summarize(id_count = n()) %>% 
  pull(id_count)

# Add id_count to startup
startup_country <- startup %>% 
  mutate(id_count) %>% 
  separate_rows(`Investor's_Origin`, sep = ", ") %>% 
  mutate(new_value = `Deal_Value_(USD)` / id_count) %>% 
  mutate(country_name = gsub(pattern = "\r\n", replacement = "", x = .$`Investor's_Origin`)) %>% 
  filter(country_name != "NA")

# Categorizing investors as foreign and local
# Type of investors (local/foreign) as two subgroups is obtained and used for monthly investment deals.
local_foreign <- startup_country %>% 
  group_by(Announcement_Date) %>% 
  summarize(local = sum(new_value[country_name == "Turkey"], na.rm = TRUE), foreign = sum(new_value[country_name != "Turkey"], na.rm = TRUE))

local_foreign %>% 
  pivot_longer(cols = c(local, foreign),names_to = "type", values_to = "deal_value") %>% 
  group_by(type) %>% 
  summarize(Sum = sum(deal_value)) %>% 
  print()

local_foreign %>% 
  pivot_longer(cols = c(local, foreign),names_to = "type", values_to = "deal_value") %>% 
  ggplot(aes(x = type, y = deal_value)) + 
  geom_col(stat="identity", fill="#69b3a2") +  
  scale_y_continuous(labels = comma_format(big.mark = ".",
                                           decimal.mark = ",")) +
  labs(x = "Investor Type", y = "Total Value of Deals", title = "Domestic vs. Foreign Investment") +
  theme_bw()
```



### Analysing Country Profile of Investors

Although number of deals accomplished by local investors is higher than foreigners, total value of transactions completed by foreign investors is 12 times higher than locals.

```{r warning=FALSE}
startup_country_graph <- startup_country %>%   
  group_by(country_name) %>% 
  summarise(country_total = sum(new_value, na.rm = TRUE), country_number = n()) %>%
  arrange(desc(country_total)) %>% 
  print()

startup_country_graph$country_name[startup_country_graph$country_name == "UAE"] = "United Arab Emirates"

# Graph investment deals and number of transactions of highest 10 countries in 2021
startup_country_graph %>% 
  arrange(country_total) %>% 
  mutate(Country=factor(country_name, country_name)) %>%
  top_n(n = 10, wt = country_total) %>% 
  ggplot(aes(x = Country, y = country_total)) +
  geom_col(stat = "identity", fill="#69b3a2") + 
  coord_flip() +  
  scale_y_continuous(labels = comma_format(big.mark = ".",
                                           decimal.mark = ",")) +
  labs(y = "Total Value of Deals", title = "Value of Startup Deals in 2021 (per Investor's Origin Country)") +
  theme_bw()

startup_country_graph %>% 
  arrange(country_number) %>% 
  mutate(Country=factor(country_name, country_name)) %>%
  top_n(n = 10, wt = country_number) %>% 
  ggplot(aes(x = Country, y = country_number)) +
  geom_col(stat="identity", fill="#69b3a2") + 
  coord_flip() +
  labs(y = "Number of Deals", title = "Number of Startup Deals in 2021 (per Investor's Origin Country)") +
  theme_bw()

## Alternative way to demonstrate geographic variation of investments into startups

world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")

# World heat map plot
ggplot(startup_country_graph) +
  geom_map(
    dat = world_map, map = world_map, aes(map_id = region),
    fill = "white", color = "#7f7f7f", size = 0.25
  ) +
  geom_map(map = world_map, aes(map_id = country_name, fill = country_total/1e6), size = 0.25) +
  scale_fill_gradient(low = "#56B1F7", high = "#132B43", name = "Investments\n(Million USD)") +
  expand_limits(x = world_map$long, y = world_map$lat) + labs(title = "Geographic Distribution of Startup Investments in 2021") +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank())
```

