---
title: "Group Project EDA"
editor: visual
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Introduction

### Load Necessary Packages for EDA

```{r library, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(readxl)
library(ggridges)
library(scales)
library(gt)
library(rmarkdown)
library(quarto)
library(knitr)
library(data.table)
library(ggthemes)
library(viridis)
```

### Load Necessary Datasets for EDA

```{r datasets, message=FALSE, warning=FALSE}

bs <- readRDS("docs/Project_Data/balance_sheet.rds")
i_s <- readRDS("docs/Project_Data/income_statement.rds")
trade_df <- readRDS("docs/Project_Data/trade.rds")
price_df <- readRDS("docs/Project_Data/input_cost.rds")
gdp_df <- readRDS("docs/Project_Data/gdp.rds")

```

### Few Data Manipulation Steps to Prepare Data for Final Stage

```{r data_preparation, message=FALSE, warning=FALSE}

bs$accounts <- bs$accounts %>% gsub("\\s+", "_", .)

bs_df <- bs %>% 
  filter(accounts %in% c("I-Current_assets", 
                         "III-Short-term_liabilities", 
                         "IV-Long-term_liabilities", 
                         "Total_assets", 
                         "V-Own_funds")) %>%
  pivot_wider(names_from = accounts, values_from = value)

i_s$accounts <- i_s$accounts %>% gsub("\\s+", "_", .)

net_income <- i_s %>% 
  filter(accounts == "Net_profit_or_loss_for_the_financial_year") %>% 
  pivot_wider(names_from = accounts, values_from = value) %>% 
  pull(Net_profit_or_loss_for_the_financial_year)
  
bs_df <- bs_df %>% mutate(net_income_is = net_income)


```

## Financial Indicators

Liquidity:

$Current\,Ratio = \frac{Current\,Assets}{Short\,Term\,Liabilities}$

Leverage:

$Leverage = \frac{Total\,Loans}{Total\,Assets}$

Profitability:

$Return\,on\,Assets = \frac{Net\,Profit}{Total\,Assets}$

```{r Financial_Indicators, message=FALSE, warning=FALSE}

bs_df_new <- bs_df %>% 
  mutate(current_ratio = bs_df$"I-Current_assets" / bs_df$"III-Short-term_liabilities", 
         leverage = bs_df$"III-Short-term_liabilities" / bs_df$"Total_assets",
         return_on_equity = bs_df$"net_income_is" / bs_df$"V-Own_funds",
         return_on_assets = bs_df$"net_income_is" / bs_df$"Total_assets") %>% 
  select(c(sector_code:year),c(current_ratio:return_on_assets))

bs_df_new <- bs_df_new %>% 
  pivot_longer(cols = c("current_ratio", "leverage", "return_on_equity", "return_on_assets"), 
               names_to = "Indicator", 
               values_to = "Value")

```

### Basic Financial Indicators of Turkish Economy

```{r economy, message=FALSE, warning=FALSE}

#Calculation and plot of weighted average financial indicators
bs_df_economy <- bs_df %>% 
  mutate(current_ratio = bs_df$"I-Current_assets" / bs_df$"III-Short-term_liabilities", 
         leverage = bs_df$"III-Short-term_liabilities" / bs_df$"Total_assets",
         return_on_equity = bs_df$"net_income_is" / bs_df$"V-Own_funds",
         return_on_assets = bs_df$"net_income_is" / bs_df$"Total_assets") %>% 
  select(c(sector_code:year),c(current_ratio:return_on_assets), Total_assets)


bs_df_economy_graph <- bs_df_economy %>% 
  group_by(year) %>% 
  summarize(Liquidity = weighted.mean(current_ratio, Total_assets) * 100,
            Leverage = weighted.mean(leverage, Total_assets) * 100,
            Profitability = weighted.mean(return_on_assets, Total_assets) * 100)

bs_df_economy_graph %>% 
  pivot_longer(-year, names_to = "indicator_yearly", values_to = "value") %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~indicator_yearly, scales = "free") +
  theme_bw() +
  labs(y = "Levels", title = "Weighted Average of Financial Indicators")

#Calculation and plot of standard deviation of financial indicators
bs_df_economy %>% 
  group_by(year) %>% 
  summarize(Liquidity = sd(current_ratio) * 10,
            Leverage = sd(leverage) * 10,
            Profitability = sd(return_on_assets) * 10) %>% 
  pivot_longer(-year, names_to = "std_yearly", values_to = "value") %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~std_yearly, scales = "free") +
  theme_bw()+
  labs(y = "Levels", title = "Standard Deviation of Financial Indicators")

```

### Table of financial indicators for the economy from 2009 to 2021

```{r table_fi}

colnames(bs_df_economy_graph) <- c("Year", "Liquidity", "Leverage", "Profitability")
bs_df_economy_graph[,2:4] <- lapply(bs_df_economy_graph[,2:4], 
                                    FUN = function(x) round(x, 2))
kable(bs_df_economy_graph)
```

### Time Series of Financial Indicators for each sector

```{r time_series_fi}

ggplot(data = bs_df_new[bs_df_new$Indicator == "current_ratio",], aes(x = year, y = Value)) +
  geom_line() +
  labs(title = "Liquidity Indicator of Real Sector Companies from 2009 to 2021", x = "Year") +
  facet_wrap(~sector_name, scales = "free") +
  theme(panel.grid = element_blank()) +
  theme_bw()


ggplot(data = bs_df_new[bs_df_new$Indicator == "leverage",], aes(x = year, y = Value)) +
  geom_line() +
  labs(title = "Leverage Indicator of Real Sector Companies from 2009 to 2021", x = "Year") +
  facet_wrap(~sector_name, scales = "free") +
  theme(panel.grid = element_blank()) +
  theme_bw()


ggplot(data = bs_df_new[bs_df_new$Indicator == "return_on_assets",], aes(x = year, y = Value)) +
  geom_line() +
  labs(title = "Profitability (ROA) Indicator of Real Sector Companies from 2009 to 2021", x = "Year") +
  facet_wrap(~sector_name, scales = "free") +
  theme(panel.grid = element_blank()) +
  theme_bw()


```

### Scatter Plots of Financial Indicators

```{r scatter_fi}

bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  ggplot(aes(x = leverage, y = current_ratio, color = sector_name)) +
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  labs(title = "Liquidity vs Leverage", y = "Liquidity") +
  theme_bw()

bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  ggplot(aes(x = current_ratio, y = return_on_assets, color = sector_name)) +
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  labs(title = "Profitability vs Liquidity", y = "Profitability", x = "Liquidity") +
  theme_bw()

bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  ggplot(aes(x = leverage, y = return_on_assets, color = sector_name)) +
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE) + 
  labs(title = "Profitability vs Leverage", y = "Profitability") +
  theme_bw()

```

### 2009 - 2019 Period Average of Financial Indicators

Liquidity Indicator

```{r liq_ave}
bs_df_average <-  bs_df_new %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year %in% 2009:2019) %>% 
  group_by(sector_name) %>% 
  summarise(liq_ave = mean(current_ratio), lev_ave = mean(leverage), pro_ave = mean(return_on_assets))

bs_df_new %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year %in% 2020:2021) %>% 
  select(c(sector_name, year, current_ratio)) %>% 
  pivot_wider(names_from = year, values_from = current_ratio) %>% 
  left_join(bs_df_average[,c("sector_name", "liq_ave")], by = "sector_name") %>% 
  rename("2009_2019_average" = "liq_ave") %>% 
  pivot_longer(-sector_name, names_to = "period") %>% 
  ggplot(aes(x = sector_name, y = value, fill = factor(period))) +
  geom_col(position = "dodge2")+
  coord_flip() +
  labs(y = "Current Ratio", x = "Sectors", title = "Liquidity Indicator", fill = "Years") +
  theme_bw()
```

Leverage Indicator

```{r lev_ave}
bs_df_new %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year %in% 2020:2021) %>% 
  select(c(sector_name, year, leverage)) %>% 
  pivot_wider(names_from = year, values_from = leverage) %>% 
  left_join(bs_df_average[,c("sector_name", "lev_ave")], by = "sector_name") %>% 
  rename("2009_2019_average" = "lev_ave") %>% 
  pivot_longer(-sector_name, names_to = "period") %>% 
  ggplot(aes(x = sector_name, y = value, fill = factor(period))) +
  geom_col(position = "dodge2")+
  coord_flip() +
  labs(y = "Leverage", x = "Sectors", title = "Leverage Indicator", fill = "Years") + theme_bw() 
```

Profitability Indicator

```{r pro_ave}
bs_df_new %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year %in% 2020:2021) %>% 
  select(c(sector_name, year, return_on_assets)) %>% 
  pivot_wider(names_from = year, values_from = return_on_assets) %>% 
  left_join(bs_df_average[,c("sector_name", "pro_ave")], by = "sector_name") %>% 
  rename("2009_2019_average" = "pro_ave") %>% 
  pivot_longer(-sector_name, names_to = "period") %>% 
  ggplot(aes(x = sector_name, y = value, fill = factor(period))) +
  geom_col(position = "dodge2")+
  coord_flip() +
  labs(y = "Profitability", x = "Sectors", title = "Return on Assets", fill = "Years") +
  theme_bw()
```

### Density Plots of Financial Indicators

```{r}
# Liquidity
bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  select(year, current_ratio) %>% 
  arrange(year) %>% 
  ggplot(aes(x = current_ratio, y = factor(year), fill = stat(quantile))) +
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient", rel_min_height = 0.005) +
  scale_fill_brewer(name = "") +
  labs(y = "Year", x = "Liquidity")

# Leverage
bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  select(year, leverage) %>% 
  arrange(year) %>% 
  ggplot(aes(x = leverage, y = factor(year), fill = stat(quantile))) +
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient", rel_min_height = 0.005) +
  scale_fill_brewer(name = "") +
  labs(y = "Year", x = "Leverage")

# Profitability
bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  select(year, return_on_assets) %>% 
  arrange(year) %>% 
  ggplot(aes(x = return_on_assets, y = factor(year), fill = stat(quantile))) +
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient", rel_min_height = 0.005) +
  scale_fill_brewer(name = "") + 
  labs(y = "Year", x = "Return on Assets")

```

## Financial Fragility Index

```{r}

bs_df_ffi <- bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  group_by(year) %>% 
  mutate(liq_rank = percent_rank(current_ratio*-1), 
         lev_rank = percent_rank(leverage), 
         profit_rank = percent_rank(return_on_assets*-1)) %>% 
  mutate(fin_fra_index = (liq_rank + lev_rank + profit_rank)/3) %>% 
  select(-c(current_ratio:return_on_assets))

bs_df_ffi %>% 
  select(c(sector_name, year, fin_fra_index)) %>% 
  pivot_longer(-c(sector_name, year), names_to = "index", values_to = "value") %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line() + 
  facet_wrap(~sector_name, scales = "free") + 
  labs(y = "Financial Fragility Index") +
  theme_bw()

```

### Heatmap for 2021

```{r fi_heatmap}
heatmap_df <- bs_df_economy %>% 
  filter(year == 2021) %>% 
  select(sector_name, current_ratio, leverage, return_on_assets) %>% 
  mutate(ffi = bs_df_ffi %>% select(year, sector_name, fin_fra_index) %>% filter(year == 2021) %>% pull(fin_fra_index)) %>% 
  arrange(desc(ffi)) 

colnames(heatmap_df) <- c("Sectors", "Liquidity", "Leverage", "Profitability", "Financial Fragility Index")

heatmap_df[,2:5] <- lapply(heatmap_df[,2:5], FUN = function(x) round(x*100, 2))

heatmap_df %>% 
  gt() %>% 
  data_color(columns = 'Liquidity', 
             colors = col_numeric(palette = c("red", "white", "green"), 
                                  domain = c(min(heatmap_df[,2]), max(heatmap_df[,2])))) %>% 
  data_color(columns = 'Leverage', 
             colors = col_numeric(palette = c("green", "white", "red"), 
                                  domain = c(min(heatmap_df[,3]), max(heatmap_df[,3])))) %>% 
  data_color(columns = 'Profitability', 
             colors = col_numeric(palette = c("red", "white", "green"), 
                                  domain = c(min(heatmap_df[,4]), max(heatmap_df[,4])))) %>% 
  data_color(columns = 'Financial Fragility Index', 
             colors = col_numeric(palette = c("green", "white", "red"), 
                                  domain = c(min(heatmap_df[,5]), max(heatmap_df[,5]))))
```

### International Trade

```{r trade}
trade_df_new <- 
  trade_df %>% 
  mutate(net_export = (export - import)/1e6) %>% 
  filter(year != 2022) %>% 
  select(sector_name, year, net_export)

df_fin <- bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value)


trade_df_new %>% 
  ggplot(aes(x = year, y = net_export))+
  geom_line() +
  facet_wrap(~sector_name, scales = "free")
  

trade_df_new %>% 
  group_by(sector_name) %>% 
  summarize(average = round(mean(net_export), 2)) %>% 
  arrange(average) %>% 
  kable(caption = "Average Net Exports")

trade_df_new %>% 
  filter(sector_name %in% c("Agriculture", "Manufacturing", "Mining", "Water_Supply")) %>% 
  inner_join(df_fin %>% select(-sector_code, -group), by = c("sector_name", "year")) %>% 
  ggplot(aes(x = net_export, y = return_on_assets))+
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE) + 
  facet_wrap(~sector_name, scales = "free")

trade_df_new %>% 
  filter(sector_name %in% c("Agriculture", "Manufacturing", "Mining", "Water_Supply")) %>% 
  inner_join(df_fin %>% select(-sector_code, -group), by = c("sector_name", "year")) %>% inner_join(bs_df_ffi %>% select(sector_name, year, fin_fra_index), by = c("sector_name", "year")) %>% 
  ggplot(aes(x = net_export, y = fin_fra_index))+
  geom_point() + 
  geom_smooth(method='lm', formula= y~x) + 
  facet_wrap(~sector_name, scales = "free")
```

### Tradable vs Non-Tradable Sectors

```{r trade_split}

tradables <- c("Agriculture", "Manufacturing", "Mining", "Water_Supply")

df_fin <- df_fin %>% inner_join(bs_df_ffi %>% select(sector_name, year, fin_fra_index), by = c("sector_name", "year"))

df_fin %>% 
  select(-sector_code, -group) %>% 
  mutate(trade_group = ifelse(sector_name %in% tradables, "tradable", "non_tradable")) %>% 
  group_by(year, trade_group) %>% 
  summarize(financial_fragility_index = mean(fin_fra_index)) %>% 
  ggplot(aes(year, financial_fragility_index, color = trade_group)) + 
  geom_line()

```

### Producer Price Inflation

For each sector, average increases in producer prices (aka input costs) is calculated and given in the table below.

Table shows average increases in input costs of sectors in Turkey from 2017 to 2021.

-   Top 3 sectors with highest average price increases are **Electricity-Gas**, **Manufacturing** and **Transportation** with average price increases %23.4, %20.2 and %19.5 consecutively.

-   Information sector is the one experiencing the lowest inrease in its inputs.

```{r ppi}

pct_change <- function(x){(x / lag(x, 12) -1)*100}

price_df <- price_df %>% filter(date >= as.Date("2017-01-01") & date < as.Date("2022-01-01")) %>% 
  select(-sector_code, -group) %>% 
  pivot_wider(names_from = sector_name, values_from = index)

price_change_df <- cbind(date = price_df$date, sapply(price_df[,2 : 11], pct_change) %>% as_tibble())
  
year_end_inf <- price_change_df %>% 
  filter(!is.na(Mining) & month(date) == 1) %>% 
  pivot_longer(-date, names_to = "sector_name", values_to = "inflation") 

year_end_inf$date <- year(year_end_inf$date)

year_end_inf <- rename(year_end_inf, "year" = "date")

#Average Inflation for the period 2018-2021
table_inflation <- year_end_inf %>% 
  group_by(sector_name) %>% 
  summarize(average_inflation = round(mean(inflation),2)) %>% 
  arrange(desc(average_inflation))

colnames(table_inflation) <- c("Sector Name", "Inflation (%, YoY)")

kable(table_inflation, caption = "Average Increases in Input Costs")
```

### Financial Indicators with Inflation data

```{r fi-ppi}

inner_join(bs_df_new %>% 
  select(-sector_code, -group) %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year >= 2018), year_end_inf, by = c("sector_name", "year")) %>%
  filter(inflation < 50) %>% # Omit outlier observation
  ggplot(aes(x = inflation, y = return_on_assets)) + 
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)

```

### Growth Dynamics

```{r growth}

gdp_df$pct_change <- (gdp_df$gdp_index / lag(gdp_df$gdp_index, 1) - 1) * 100

gdp_df <- gdp_df %>% filter(!is.na(pct_change) & year >= 2009)

# Real GDP Growth Summary Table 

gdp_df_tb <- gdp_df

setDT(gdp_df)
gdp_df[, as.list(summary(pct_change)), by = sector]


gdp_df_tb$sector <- gdp_df_tb %>% 
  select(-gdp_index) %>% .$sector %>% gsub("-.*", "", .)


bs_df_new <- bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  left_join(bs_df_ffi %>% select(sector_code, year, fin_fra_index), by = c("sector_code", "year")) %>% 
  pivot_longer(-c(sector_code : year), names_to = "Indicator", values_to = "Value")

fin_ind_group <- bs_df_new %>% 
  select(sector_code, year, Indicator, Value) %>% 
  mutate(group = case_when(sector_code == "A" ~ "A", 
            sector_code %in% c("B", "C", "D", "E") ~ "BCDE", 
            sector_code == "F" ~ "F",
            sector_code %in% c("G", "H", "I") ~ "GHI",
            sector_code == "J" ~ "J",
            sector_code == "L" ~ "L",
            sector_code %in% c("M", "N") ~ "MN",
            TRUE ~ "other")) %>% 
  filter(group != "other") %>% 
  left_join(bs_df %>% select(sector_code, year, Total_assets), Total_assets, by = c("sector_code", "year")) %>%
  select(-sector_code) %>% 
  group_by(group, year, Indicator) %>% 
  summarise(value = weighted.mean(Value, Total_assets)) %>% pivot_wider(names_from = Indicator, values_from = value)

header <- c("A", "BCDE", "F", "GHI", "J", "L", "MN")
names <- c("Agriculture", "Industry", "Construction", "Services", "Information", "Real Estate", "Other Services")
growth_info <- data.frame(header, names)

growth_df <- left_join(fin_ind_group, gdp_df_tb %>% select(-gdp_index),by = c("group" = "sector", "year"))
growth_df <- growth_df %>% 
  left_join(growth_info, by = c("group" = "header"))

growth_df %>% 
  ggplot(aes(lag(pct_change,1), fin_fra_index)) +
  geom_point() + 
  facet_wrap(~names,  scales = "free") + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)
  
```

## Sectoral Gender Distribution

```{r}
#Gender distribution in main sectors that are agriculture, construction, industry and services.
emp_data <- readRDS("docs/Project_Data/employment.rds")

table <- emp_data %>%
  group_by(group, gender) %>%
  arrange(sector_name) %>%
  mutate("value") %>%
  summarise(value=sum(number))

table

ggplot(table, aes(x=group, y=value, fill=gender)) +
    geom_bar(stat='identity', position='dodge')



```

```{r}
#Detailed gender distribution
emp_data <- readRDS("docs/Project_Data/employment.rds")

table_2 <- emp_data %>%
  group_by(sector_name, gender) %>%
  arrange(sector_name) %>%
  mutate("value") %>%
  summarise(value=sum(number))

ggplot(table_2, aes(x=sector_name, y=value, fill=gender)) +
    geom_bar(stat='identity', position='dodge') + 
    theme(axis.text.x = element_text(angle=90, hjust=1))

```

```{r}
#Percentage of the number of employees in the sectors
emp_data <- readRDS("docs/Project_Data/employment.rds")

table_3 <- emp_data %>%
  group_by(sector_name) %>%
  arrange(sector_name) %>%
  mutate("value") %>%
  summarise(value=sum(number))

total = sum(table_3$value)

table_4 <- table_3 %>%
  group_by(sector_name) %>%
  arrange(sector_name) %>%
  mutate("percentage") %>%
  summarise(percentage=100*value/total)

table_4

```

## Firm Size Comparison by Sector Groups

```{r}
#Firm size comparison in the sectors; agriculture, construction, industry and services.
options(dplyr.summarise.inform = FALSE)
sectorinf_data <- readRDS("docs/Project_Data/sector_information.rds")

table_sectorinf <- sectorinf_data %>%
  filter(number == 'Number_of_companies') %>%
  filter(size != 'Total') %>%
  group_by(group,size) %>%
  mutate("num_of_companies") %>%
  summarise(num_of_companies=sum(value)) %>%
  arrange(group,desc(size))

table_sectorinf

table_sectorinf %>%
  ggplot(aes(x=group,y=num_of_companies, fill=size)) + 
  geom_bar(stat='identity', position='dodge') +
  xlab("Number of companies") +
  ylab("Sector group")
```

```{r}
#Firm size comparison by sector names as a detailed version.
table_sectordetail <- sectorinf_data %>%
  filter(number == 'Number_of_companies') %>%
  filter(size != 'Total') %>%
  group_by(sector_name, size) %>%
  mutate("num_of_companies") %>%
  summarise(num_of_companies=sum(value)) %>%
  arrange(sector_name,desc(size))

table_sectordetail

table_sectordetail %>%
  ggplot(aes(y=sector_name,x=num_of_companies, fill=size)) + 
  geom_bar(stat='identity', position='dodge') +
  xlab("Number of companies") +
  ylab("Detailed sector group")
```
