---
title: "Group Project EDA"
editor: visual
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Introduction

### Load Necessary Packages for EDA

```{r library, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(readxl)
library(ggridges)
library(scales)
library(gt)
library(rmarkdown)
library(quarto)
library(knitr)
library(data.table)
```

### Load Necessary Datasets for EDA

```{r datasets, message=FALSE, warning=FALSE}

bs <- readRDS("docs/Project_Data/balance_sheet.rds")
i_s <- readRDS("docs/Project_Data/income_statement.rds")
trade_df <- readRDS("docs/Project_Data/trade.rds")
price_df <- readRDS("docs/Project_Data/input_cost.rds")
gdp_df <- readRDS("docs/Project_Data/gdp.rds")

```

### Few Data Manipulation Steps to Prepare Data for Final Stage

```{r data_preparation, message=FALSE, warning=FALSE}

bs$accounts <- bs$accounts %>% gsub("\\s+", "_", .)

bs_df <- bs %>% 
  filter(accounts %in% c("I-Current_assets", 
                         "III-Short-term_liabilities", 
                         "IV-Long-term_liabilities", 
                         "Total_assets", 
                         "V-Own_funds")) %>%
  pivot_wider(names_from = accounts, values_from = value)

i_s$accounts <- i_s$accounts %>% gsub("\\s+", "_", .)

net_income <- i_s %>% 
  filter(accounts == "Net_profit_or_loss_for_the_financial_year") %>% 
  pivot_wider(names_from = accounts, values_from = value) %>% 
  pull(Net_profit_or_loss_for_the_financial_year)
  
bs_df <- bs_df %>% mutate(net_income_is = net_income)


```

## Financial Indicators

```{r Financial_Indicators, message=FALSE, warning=FALSE}

bs_df_new <- bs_df %>% 
  mutate(current_ratio = bs_df$"I-Current_assets" / bs_df$"III-Short-term_liabilities", 
         leverage = bs_df$"III-Short-term_liabilities" / bs_df$"Total_assets",
         return_on_equity = bs_df$"net_income_is" / bs_df$"V-Own_funds",
         return_on_assets = bs_df$"net_income_is" / bs_df$"Total_assets") %>% 
  select(c(sector_code:year),c(current_ratio:return_on_assets))

bs_df_new <- bs_df_new %>% 
  pivot_longer(cols = c("current_ratio", "leverage", "return_on_equity", "return_on_assets"), 
               names_to = "Indicator", 
               values_to = "Value")

```

### Economy Wide Financial Indicators

```{r economy, message=FALSE, warning=FALSE}

#Calculation and plot of weighted average financial indicators
bs_df_economy <- bs_df %>% 
  mutate(current_ratio = bs_df$"I-Current_assets" / bs_df$"III-Short-term_liabilities", 
         leverage = bs_df$"III-Short-term_liabilities" / bs_df$"Total_assets",
         return_on_equity = bs_df$"net_income_is" / bs_df$"V-Own_funds",
         return_on_assets = bs_df$"net_income_is" / bs_df$"Total_assets") %>% 
  select(c(sector_code:year),c(current_ratio:return_on_assets), Total_assets)


bs_df_economy_graph <- bs_df_economy %>% 
  group_by(year) %>% 
  summarize(liq_yearly = weighted.mean(current_ratio, Total_assets),
            lev_yearly = weighted.mean(leverage, Total_assets),
            pro_yearly = weighted.mean(return_on_assets, Total_assets))

bs_df_economy_graph %>% 
  pivot_longer(-year, names_to = "indicator_yearly", values_to = "value") %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~indicator_yearly, scales = "free")

#Calculation and plot of standard deviation of financial indicators
bs_df_economy %>% 
  group_by(year) %>% 
  summarize(liq_yearly_std = sd(current_ratio),
            lev_yearly_std = sd(leverage),
            pro_yearly_std = sd(return_on_assets)) %>% 
  pivot_longer(-year, names_to = "std_yearly", values_to = "value") %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~std_yearly, scales = "free")

```

### Table of financial indicators for the economy from 2009 to 2021

```{r table_fi}
kable(bs_df_economy_graph)
```

### Time Series of Financial Indicators for each sector

```{r time_series_fi}

ggplot(data = bs_df_new[bs_df_new$Indicator == "current_ratio",], aes(x = year, y = Value)) +
  geom_line() +
  labs(title = "Liquidity Indicator of Real Sector Companies from 2009 to 2021", x = "Year") +
  facet_wrap(~sector_name, scales = "free") +
  theme(panel.grid = element_blank())


ggplot(data = bs_df_new[bs_df_new$Indicator == "leverage",], aes(x = year, y = Value)) +
  geom_line() +
  labs(title = "Leverage Indicator of Real Sector Companies from 2009 to 2021", x = "Year") +
  facet_wrap(~sector_name, scales = "free") +
  theme(panel.grid = element_blank())    


ggplot(data = bs_df_new[bs_df_new$Indicator == "return_on_assets",], aes(x = year, y = Value)) +
  geom_line() +
  labs(title = "Profitability (ROA) Indicator of Real Sector Companies from 2009 to 2021", x = "Year") +
  facet_wrap(~sector_name, scales = "free") +
  theme(panel.grid = element_blank())


```

### Scatter Plots of financial Indicators

```{r scatter_fi}

bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  ggplot(aes(x = leverage, y = current_ratio, color = sector_name)) +
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)

bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  ggplot(aes(x = current_ratio, y = return_on_assets, color = sector_name)) +
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)

bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  ggplot(aes(x = leverage, y = return_on_assets, color = sector_name)) +
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)

```

### 2009 - 2019 Period Average of Financial Indicators
```{r average_fi}
bs_df_average <-  bs_df_new %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year %in% 2009:2019) %>% 
  group_by(sector_name) %>% 
  summarise(liq_ave = mean(current_ratio), lev_ave = mean(leverage), pro_ave = mean(return_on_assets))

bs_df_new %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year %in% 2020:2021) %>% 
  select(c(sector_name, year, current_ratio)) %>% 
  pivot_wider(names_from = year, values_from = current_ratio) %>% 
  left_join(bs_df_average[,c("sector_name", "liq_ave")], by = "sector_name") %>% 
  rename("2009_2019_average" = "liq_ave") %>% 
  pivot_longer(-sector_name, names_to = "period") %>% 
  ggplot(aes(x = sector_name, y = value, fill = factor(period))) +
  geom_col(position = "dodge2")
```

### Density Plots of Financial Indicators

```{r}
# Liquidity
bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  select(year, current_ratio) %>% 
  arrange(year) %>% 
  ggplot(aes(x = current_ratio, y = factor(year), fill = stat(quantile))) +
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient", rel_min_height = 0.005) +
  scale_fill_brewer(name = "")

# Leverage
bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  select(year, leverage) %>% 
  arrange(year) %>% 
  ggplot(aes(x = leverage, y = factor(year), fill = stat(quantile))) +
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient", rel_min_height = 0.005) +
  scale_fill_brewer(name = "")

# Profitability
bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  select(year, return_on_assets) %>% 
  arrange(year) %>% 
  ggplot(aes(x = return_on_assets, y = factor(year), fill = stat(quantile))) +
  stat_density_ridges(quantile_lines = FALSE,
                      calc_ecdf = TRUE,
                      geom = "density_ridges_gradient", rel_min_height = 0.005) +
  scale_fill_brewer(name = "")

```



## Financial Fragility Index

```{r}

bs_df_ffi <- bs_df_new %>% 
  pivot_wider(names_from = "Indicator", values_from = "Value") %>% 
  group_by(year) %>% 
  mutate(liq_rank = percent_rank(current_ratio*-1), 
         lev_rank = percent_rank(leverage), 
         profit_rank = percent_rank(return_on_assets*-1)) %>% 
  mutate(fin_fra_index = (liq_rank + lev_rank + profit_rank)/3) %>% 
  select(-c(current_ratio:return_on_assets))

bs_df_ffi %>% 
  select(c(sector_name, year, fin_fra_index)) %>% 
  pivot_longer(-c(sector_name, year), names_to = "index", values_to = "value") %>% 
  ggplot(aes(x = year, y = value, color = sector_name)) +
  geom_line() + 
  facet_wrap(~sector_name, scales = "free")

```
### Heatmap for 2021

```{r fi_heatmap}
bs_df_economy %>% 
  filter(year == 2021) %>% 
  select(sector_name, current_ratio, leverage, return_on_assets) %>% 
  mutate(ffi = bs_df_ffi %>% select(year, sector_name, fin_fra_index) %>% filter(year == 2021) %>% pull(fin_fra_index)) %>% 
  arrange(desc(ffi)) %>% 
  gt() %>% 
  data_color(columns = current_ratio, 
             colors = col_numeric(palette = c("red", "white", "green"), 
                                  domain = c(min(current_ratio), max(current_ratio)))) %>% 
  data_color(columns = leverage, 
             colors = col_numeric(palette = c("green", "white", "red"), 
                                  domain = c(min(leverage), max(leverage)))) %>% 
  data_color(columns = return_on_assets, 
             colors = col_numeric(palette = c("red", "white", "green"), 
                                  domain = c(min(return_on_assets), max(return_on_assets)))) %>% 
  data_color(columns = ffi, 
             colors = col_numeric(palette = c("green", "white", "red"), 
                                  domain = c(min(ffi), max(ffi))))
```


### International Trade

```{r}
trade_df_new <- 
  trade_df %>% 
  mutate(net_export = (export - import)/1e6) %>% 
  filter(year != 2022) %>% 
  select(sector_name, year, net_export)

df_fin <- bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value)


trade_df_new %>% 
  ggplot(aes(x = year, y = net_export))+
  geom_line() +
  facet_wrap(~sector_name, scales = "free")
  

trade_df_new %>% 
  group_by(sector_name) %>% 
  summarize(mean = mean(net_export)) %>% 
  arrange(mean)

trade_df_new %>% 
  filter(sector_name %in% c("Agriculture", "Manufacturing", "Mining", "Water_Supply")) %>% 
  inner_join(df_fin %>% select(-sector_code, -group), by = c("sector_name", "year")) %>% 
  ggplot(aes(x = net_export, y = return_on_assets))+
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE) + 
  facet_wrap(~sector_name, scales = "free")

trade_df_new %>% 
  filter(sector_name %in% c("Agriculture", "Manufacturing", "Mining", "Water_Supply")) %>% 
  inner_join(df_fin %>% select(-sector_code, -group), by = c("sector_name", "year")) %>% 
  inner_join(bs_df_ffi %>% select(sector_name, year, fin_fra_index), by = c("sector_name","year")) %>% 
  ggplot(aes(x = net_export, y = fin_fra_index))+
  geom_point() + 
  geom_smooth(method='lm', formula= y~x) + 
  facet_wrap(~sector_name, scales = "free")
```

### Tradable vs Non-Tradable Sectors 

```{r trade_split}

tradables <- c("Agriculture", "Manufacturing", "Mining", "Water_Supply")
df_fin
df_fin %>% 
  select(-sector_code, -group) %>% 
  inner_join(bs_df_ffi %>% select(sector_name, year, fin_fra_index), 
             by = c("sector_name","year")) %>% 
  mutate(trade_group = ifelse(sector_name %in% tradables, "tradable", "non_tradable")) %>% 
  group_by(year, trade_group) %>% 
  summarize(financial_fragility_index = mean(fin_fra_index)) %>% 
  ggplot(aes(year, financial_fragility_index, color = trade_group)) + 
  geom_line()

```

### Producer Price Inflation

```{r ppi}

pct_change <- function(x){(x / lag(x, 12) -1)*100}

price_df <- price_df %>% filter(date >= as.Date("2017-01-01") & date < as.Date("2022-01-01")) %>% 
  select(-sector_code, -group) %>% 
  pivot_wider(names_from = sector_name, values_from = index)

price_change_df <- cbind(date = price_df$date, sapply(price_df[,2 : 11], pct_change) %>% as_tibble())
  
year_end_inf <- price_change_df %>% 
  filter(!is.na(Mining) & month(date) == 1) %>% 
  pivot_longer(-date, names_to = "sector_name", values_to = "inflation") 

year_end_inf$date <- year(year_end_inf$date)

year_end_inf <- rename(year_end_inf, "year" = "date")

#Average Inflation for the period 2018-2021
year_end_inf %>% 
  group_by(sector_name) %>% 
  summarize(average_inflation = mean(inflation)) %>% 
  arrange(desc(average_inflation))

```

### Financial Indicators with Inflation data

```{r fi-ppi}

inner_join(bs_df_new %>% 
  select(-sector_code, -group) %>% 
  pivot_wider(names_from = Indicator, values_from = Value) %>% 
  filter(year >= 2018), year_end_inf, by = c("sector_name", "year")) %>%
  filter(inflation < 50) %>% # Omit outlier observation
  ggplot(aes(x = inflation, y = return_on_assets)) + 
  geom_point() + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)

```

### Growth Dynamics

```{r growth}

gdp_df$pct_change <- (gdp_df$gdp_index / lag(gdp_df$gdp_index, 1) - 1) * 100

gdp_df <- gdp_df %>% filter(!is.na(pct_change) & year >= 2009)

# Real GDP Growth Summary Table 

gdp_df_tb <- gdp_df

setDT(gdp_df)
gdp_df[, as.list(summary(pct_change)), by = sector]


gdp_df_tb$sector <- gdp_df_tb %>% 
  select(-gdp_index) %>% .$sector %>% gsub("-.*", "", .)


bs_df_new <- bs_df_new %>% pivot_wider(names_from = Indicator, values_from = Value) %>% 
  left_join(bs_df_ffi %>% select(sector_code, year, fin_fra_index), by = c("sector_code", "year")) %>% 
  pivot_longer(-c(sector_code : year), names_to = "Indicator", values_to = "Value")

fin_ind_group <- bs_df_new %>% 
  select(sector_code, year, Indicator, Value) %>% 
  mutate(group = case_when(sector_code == "A" ~ "A", 
            sector_code %in% c("B", "C", "D", "E") ~ "BCDE", 
            sector_code == "F" ~ "F",
            sector_code %in% c("G", "H", "I") ~ "GHI",
            sector_code == "J" ~ "J",
            sector_code == "L" ~ "L",
            sector_code %in% c("M", "N") ~ "MN",
            TRUE ~ "other")) %>% 
  filter(group != "other") %>% 
  left_join(bs_df %>% select(sector_code, year, Total_assets), Total_assets, by = c("sector_code", "year")) %>%
  select(-sector_code) %>% 
  group_by(group, year, Indicator) %>% 
  summarise(value = weighted.mean(Value, Total_assets)) %>% pivot_wider(names_from = Indicator, values_from = value)

header <- c("A", "BCDE", "F", "GHI", "J", "L", "MN")
names <- c("Agriculture", "Industry", "Construction", "Services", "Information", "Real Estate", "Other Services")
growth_info <- data.frame(header, names)

growth_df <- left_join(fin_ind_group, gdp_df_tb %>% select(-gdp_index),by = c("group" = "sector", "year"))
growth_df <- growth_df %>% 
  left_join(growth_info, by = c("group" = "header"))

growth_df %>% 
  ggplot(aes(lag(pct_change,1), fin_fra_index)) +
  geom_point() + 
  facet_wrap(~names,  scales = "free") + 
  geom_smooth(method='lm', formula= y~x, se = FALSE)
  
```

